\chapter*{Abstract}
\addcontentsline{toc}{chapter}{ABSTRACT}

\begin{center}
	\center
	\begin{onehalfspace}
		\Large \bfseries \MakeUppercase{\thetitleinenglish}

		\normalfont \normalsize
		By

		\theauthor
	\end{onehalfspace}
\end{center}

\begin{singlespace}
	The research ``Development of an Autonomous System using Artificial
	Intelligence for Tram" focuses on developing an autonomous system for
	trams. This autonomous system's development requires simulation to
	facilitate its implementation, and the simulation uses a
	"software-in-the-loop" scheme with the CARLA simulator as the software tool.
	The simulation needs to be as close to the real world as possible to support
	the system's development. This final project implements tram, local objects,
	and typical Indonesian environments in the CARLA simulator for autonomous
	tram simulation. After exploring the CARLA editor and its documentation, 3D
	model assets are created, those assets are then imported, the imported
	assets are edited, the edited assets are integrated into the simulation, and
	implementations are then validated. The implemented objects include trams,
	public transport vehicles (``angkot"), traditional bicycles (``sepeda
	onthel"), motorcycles, pedicabs (``becak"), train stations, traffic signs,
	and railway tracks. The results of the implementation are validated and
	evaluated to ensure that these assets or objects, particularly the vehicles,
	remain stable within the simulation and can follow both user control and
	autopilot control. The implementation results show that trams and angkots
	are stable as four-wheeled vehicles and can follow user control and
	autopilot control. On the other hand, traditional bicycles, motorcycles, and
	pedicabs remain stable as static objects and are not implemented as
	two-wheeled vehicles. Train stations, traffic signs, and railway tracks are
	stable as static objects as well. Sepeda onthel, motorcycles, and pedicabs
	are not implemented as two-wheeled vehicles because they are unstable during
	the simulation. This instability is caused by asset misconfiguration. The
	implemented assets are then exported, making them available for other
	autonomous system developers to use these assets.

	Keywords: autonomous tram, simulation, CARLA simulator, Indonesia local
	objects.
\end{singlespace}

\clearpage
